---
title: "SWI mapping"
author: "Kelly Hondula"
output:
  html_document:
    toc: true
    toc_depth: 3
---

This is a test for mapping the spatial extent of salt water intrusion on farmfields across Somerset, Dorchester, and Worchester counties in Maryland. This is very simple unsupervised pixel-based classification, but we could probably do supervised classification and explore using object-based classification (shapes instead of just pixel values). Once pixels are classified/identified as being the target, it is possible to calculate the area based on the resolution of the image. 

```{r, message=FALSE}
# load libraries

library(raster)
library(randomForest)
library(cluster)
library(RStoolbox)
library(sp)
library(rgdal)
library(leaflet)
library(dplyr)

```

## Data

* [Imagery](http://imagery.geodata.md.gov:8080/ExpressZip) - includes 2013 and 2015 NAIP, 2013 and 2016 six inch CIR imagery, and 2010, 2013, and 2016 six inch imagery. 
* [MDP land use land cover interactive map](http://mdpgis.mdp.state.md.us/landuse/imap/index.html)
* [Download for 2010 LULC shapefile data by county](http://planning.maryland.gov/OurProducts/downloadFiles.shtml)

## Somerset county land use data

```{r}
# import someserset county land use data
shp <- "../data/Some_2010LULC/"
lulc <- readOGR(dsn = shp,
                    layer = "Some_2010LULC",
                    stringsAsFactors = FALSE)

# spatial transformation for plotting in leaflet
lulcT <- spTransform(lulc, CRS("+proj=longlat +datum=WGS84 +no_defs"))
# table(lulcT@data$LU_CODE)
```

[guide to `LU_CODE`](http://geodata.md.gov/imap/rest/services/PlanningCadastre/MD_LandUseLandCover/MapServer/1) shows that agriculture are codes 21, 22, 23, 25, 241, 242, 20

```{r}
lulc_AG <- lulcT[lulcT$LU_CODE %in% c(21, 22, 23, 25, 241, 242, 20),]
```

2010 agriculture land use data looks to match up pretty well with field locations from esri imagery (unsure of date). This could be used as a filter for the imagery data to mask out everything else and just use the farm fields, which would likely make the classification easier. I didn't line up the image data with this shapefile to do that yet. 

```{r}
# plot(lulc_AG)
leaflet(lulc_AG) %>%
  addProviderTiles("Esri.WorldImagery") %>%
  addPolygons()

```

## Imagery classification

Downloaded imagery data to test out classification. 

### Greenhead farm area 

CIR data - this includes color and infrared so has 4 bands instead of just 3 (red/green/blue). I thought maybe the signature of SWI areas would show up better with this, but it probably also depends on the date it was taken. 

```{r}
cir <- brick("../data/greenhead_farm/EasternShoreCIR2016/greenhead_farm.jpg")
# plot(cir)
raster::plotRGB(cir)
```

6 inch imagery data

```{r}
img <- brick("../data/greenhead_farm/EasternShore2016/greenhead_farm.jpg")
# plot(img)
raster::plotRGB(img)

```


2015 NAIP data - the white areas of SWI show up well here

```{r}
naip <- brick("../data/greenhead_farm/NAIP2015/greenhead_farm.jpg")
naip <- raster("../data/greenhead_farm/NAIP2015/greenhead_farm.jpg")
plot(boundaries(naip))
# plot(naip) # plots each of RGB raster layers independently
raster::plotRGB(naip)
```

```{r}
naip2 <- raster("../data/NAIP2015/swi.tif")
naip2 <- brick("../data/NAIP2015/swi.tif")
# plot(naip) # plots each of RGB raster layers independently
raster::plotRGB(naip2)
plot(naip2)
```

Crop to farm areas
```{r}
lulc_AG <- spTransform(lulc_AG, crs(naip2))
plot(naip2)
plot(lulc_AG, add = TRUE)

bbox(naip2)
naip_farm <- mask(naip2, mask = lulc_AG)
plot(naip_farm)
```


### Unsupervised classification

Using the `unsuperClass()` function in the RStoolbox package, with 5 classes. Since there are field validated locations, could use a supervised algorithm based on a training set of data from these locations. Could also be improved with the masking to just farm land cover areas using the LULC data above. 

```{r}
set.seed(25)
unC <- unsuperClass(naip_farm, nSamples = 100, nClasses = 5, nStarts = 5)
# unC
colors <- rainbow(5)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

```{r}
set.seed(25)
unC <- unsuperClass(naip, nSamples = 100, nClasses = 5, nStarts = 5)
# unC
colors <- rainbow(5)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

Seems like the road is getting grouped in, maybe try adding more classes. 

```{r}
set.seed(25)
unC <- unsuperClass(naip, nSamples = 100, nClasses = 6, nStarts = 5)
# unC
colors <- rainbow(6)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

10 classes

```{r}
set.seed(25)
unC <- unsuperClass(naip, nSamples = 100, nClasses = 10, nStarts = 5)
# unC
colors <- rainbow(10)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```
Try using the 6 inch imagery data

```{r}
set.seed(25)
unC <- unsuperClass(img, nSamples = 100, nClasses = 6, nStarts = 5)
# unC
colors <- rainbow(6)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```


Larger area with NAIP 2015 data
```{r}
naip2 <- brick("../data/NAIP2015/greenhead2.jpg")
raster::plotRGB(naip2)
```

unsupervised classification with 6 classes picks up the farm field edges (green) it seems like, but also the white roofs. 

```{r}
set.seed(25)
unC <- unsuperClass(naip2, nSamples = 100, nClasses = 6, nStarts = 5)
# unC
colors <- rainbow(6)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

Almodinton area

```{r}
alm <- brick("../data/almodinton/NAIP2015/almodinton_3.jpg")
raster::plotRGB(alm)
```

Classification

```{r}
set.seed(25)
unC <- unsuperClass(alm, nSamples = 100, nClasses = 6, nStarts = 5)
# unC

colors <- rainbow(6)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

10 classes
```{r}
set.seed(25)
unC <- unsuperClass(alm, nSamples = 100, nClasses = 10, nStarts = 5)
# unC

colors <- rainbow(10)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```

## Fitzgerald farm area

```{r}
fitz <- brick("../data/fitzgerald_farm/NAIP2015/fitzgerald_farm.jpg")
raster::plotRGB(fitz)
```

Classification (6)

```{r}
set.seed(25)
unC <- unsuperClass(fitz, nSamples = 100, nClasses = 6, nStarts = 5)
# unC

colors <- rainbow(6)
plot(unC$map, col = colors, legend = FALSE, axes = FALSE, box = FALSE)

```


